---
title: "DivDis_Analyses"
author: "Forest Schenck"
date: "January 29, 2018"
output: pdf_document
---

#Packages
```{r}
#data manipulation
library(dplyr)
library(tidyr)

#plotting
library(ggplot2)
library(sciplot)
library(multcompView)
```

#Uploading data
```{r}
mesocosm.data <- read.csv("ConditionAssessment_Data.csv")
```

#Data organization
```{r}
str(mesocosm.data)

#subsetting data to january 2018 variables of interest
jan2018 <- mesocosm.data %>%
  filter(month == "january") %>% #subsetting to month
  dplyr::select(bin.id, paired.com.id, vase.id, loc.id, diversity.treatment, depth, disease.treatment, family.id, color, alive, shoot.no, length.1:length.7, Notes) #only selection variables of interest

for(n in 1:480){
  jan2018$total.length[n] <- sum(jan2018[n,12:18], na.rm=T) #summing length of all shoots per plant
}

#gathering data so disease and control are on same line
jan2018.disease <- jan2018 %>%
  filter(disease.treatment == "disease") %>%
  dplyr::select(bin.id, paired.com.id, loc.id, diversity.treatment, depth, family.id, color, alive)
jan2018.control <- jan2018 %>%
  filter(disease.treatment == "control") %>%
  dplyr::select(bin.id, paired.com.id, loc.id, diversity.treatment, depth, family.id, color, alive)
colnames(jan2018.control) <- c("bin.id", "paired.com.id", "loc.id", "diversity.treatment", "depth", "family.id", "color", "alive.control")
colnames(jan2018.disease) <- c("bin.id", "paired.com.id", "loc.id", "diversity.treatment", "depth", "family.id", "color", "alive.disease")
jan2018.paired <- merge(jan2018.disease, jan2018.control, by = c("bin.id", "paired.com.id", "loc.id", "diversity.treatment", "depth", "family.id", "color"))

#VIABLE mono communities
jan2018.paired.vase.mono <- jan2018.paired %>%
  filter(diversity.treatment == "mono") %>%
  group_by(bin.id, paired.com.id, diversity.treatment, depth, family.id) %>%
  summarize(sum.alive.d = sum(alive.disease), sum.alive.c = sum(alive.control)) %>%
  filter(sum.alive.d > 1 & sum.alive.c > 1)

#MONO DATAFRAME TO EXPORT
jan2018.mono <- jan2018 %>%
  dplyr::select(bin.id, paired.com.id, vase.id, loc.id, diversity.treatment, depth, disease.treatment, family.id, color, alive) %>%
  filter(diversity.treatment == "mono") %>%
  filter(alive == 1) %>%
  filter(family.id == "S6" | family.id == "S3"| family.id == "S12"| family.id == "S11"| family.id == "S1" | family.id == "M9"| family.id == "M7"| family.id == "M6"| family.id == "M4"| family.id == "M3"| family.id == "M12"| family.id == "M10"| family.id == "D8"| family.id == "D6"| family.id == "D4"| family.id == "D3"| family.id == "D11"| family.id == "D10")

write.csv( jan2018.mono, "MonoVases.csv" )

##identifying VIABLE communities based on pairs present (must be at least 2 shoot alive in both and must be at least 2 shoots from same mom alive in both and at least 2 of surving shoots pairs must be in monoculture)
#VIABLE poly communities
jan2018.paired.poly <- jan2018.paired %>%
  filter(diversity.treatment == "poly") #only selecting polys
jan2018.paired.poly$pair.present <- jan2018.paired.poly$alive.disease == jan2018.paired.poly$alive.control #creating a column identifying if pairs are matching or not
jan2018.paired.poly <- jan2018.paired.poly %>%
  filter(alive.disease ==1 | alive.control == 1) #each pair must have at least 1 live shoot to be considered (excluded 0=0 issue)
for(n in 1:108){
  if (jan2018.paired.poly$pair.present[n] == T)
  {jan2018.paired.poly$pair.present[n] = 1
  } else {
    jan2018.paired.poly$pair.present[n] = 0
    }
} #switching pairs from T/F to 1/0

jan2018.paired.vase.poly <- jan2018.paired.poly %>%
  group_by(bin.id, paired.com.id, diversity.treatment, depth) %>% #grouping by vase
  summarize(sum.pair.present = sum(pair.present), sum.alive.d = sum(alive.disease), sum.alive.c = sum(alive.control)) %>%
  filter(sum.pair.present > 1) #only selected vases with at least 2 paired locations where shoots are alive

jan2018.paired.poly.subset.complete <- jan2018.paired.poly %>%
  filter(paired.com.id == "V19" | paired.com.id == "V21"| paired.com.id == "V23"| paired.com.id == "V25"| paired.com.id == "V55"| paired.com.id == "V59"| paired.com.id == "V61"| paired.com.id == "V71"| paired.com.id == "V91"| paired.com.id == "V93"| paired.com.id == "V95"| paired.com.id == "V97"| paired.com.id == "V101"| paired.com.id == "V103"| paired.com.id == "V107"| paired.com.id == "V111"| paired.com.id == "V117") %>% #selecting vases with a least 2 perfect pairs
  #filter(pair.present != 0) %>% #removing all shoots without a pair
  filter(family.id == "S6" | family.id == "S3"| family.id == "S12"| family.id == "S11"| family.id == "S1" | family.id == "M9"| family.id == "M7"| family.id == "M6"| family.id == "M4"| family.id == "M3"| family.id == "M12"| family.id == "M10"| family.id == "D8"| family.id == "D6"| family.id == "D4"| family.id == "D3"| family.id == "D11"| family.id == "D10") #only selecting shoots that have viable monocultures

###data set with vases that have exact matches without stragglers (perfect) and with stragglers that have viable Mono
jan2018.paired.vase.poly.subset.complete <- jan2018.paired.poly.subset.complete %>%
  group_by(bin.id, paired.com.id, diversity.treatment, depth) %>% #grouping by vase pairs
  summarize(sum.pair.present = sum(pair.present), sum.alive.c = sum(alive.control), sum.alive.d = sum(alive.disease)) %>% #summing up number of pairs of shoots present per vase pair
  filter(sum.pair.present > 1) #filtering out vases with less than 2 shoot pairs alive
jan2018.paired.vase.poly.subset.complete$unpaired.control <- jan2018.paired.vase.poly.subset.complete$sum.alive.c - jan2018.paired.vase.poly.subset.complete$sum.pair.present #identifying pairs with viable-mono-(vm)-stragglers in control
jan2018.paired.vase.poly.subset.complete$unpaired.disease <- jan2018.paired.vase.poly.subset.complete$sum.alive.d - jan2018.paired.vase.poly.subset.complete$sum.pair.present #identifying pairs with vm-stragglers in disease

jan2018.paired.vase.poly.subset.complete$em.vms <- NA #(em.vms = exact-match-viable-mono-straggler)
for(n in 1:10){
  if (jan2018.paired.vase.poly.subset.complete$unpaired.control[n] + jan2018.paired.vase.poly.subset.complete$unpaired.disease[n] > 0)
  {jan2018.paired.vase.poly.subset.complete$em.vms[n] = 1
  } else {
    jan2018.paired.vase.poly.subset.complete$em.vms[n] = 0
  }
}
jan2018.paired.vase.poly.subset.complete$em <- 1-jan2018.paired.vase.poly.subset.complete$em.vms

##Data set that identifies location of stragglers
jan2018.paired.poly.subset.complete <- jan2018.paired.poly.subset.complete %>%
filter(paired.com.id == "V93" | paired.com.id == "V91" | paired.com.id == "V107" | paired.com.id == "V95" | paired.com.id == "V55" | paired.com.id == "V71" | paired.com.id == "V59" | paired.com.id == "V25" | paired.com.id == "V21" | paired.com.id == "V23" ) #selecting only vases that have at least 2 shoot pairs alive in both disease and control treatment

#jan2018.ems <- jan2018.paired.poly %>% #identifying exact match paired stragglers that are not in poly
  #filter(paired.com.id == "V93" | paired.com.id == "V91" | paired.com.id == "V107" | paired.com.id == "V95" | paired.com.id == "V55" | #paired.com.id == "V71" | paired.com.id == "V59" | paired.com.id == "V25" | paired.com.id == "V21" | paired.com.id == "V23" ) %>%
  #group_by(bin.id, paired.com.id, diversity.treatment, depth) %>% #grouping by vase pairs
  #summarize(sum.pair.present = sum(pair.present), sum.alive.c = sum(alive.control), sum.alive.d = sum(alive.disease))  #identifying pairs with viable-mono-(vm)-stragglers in control
#jan2018.ems$unpaired.control <- jan2018.ems$sum.alive.c - jan2018.paired.vase.poly.subset.complete$sum.pair.present 
#jan2018.paired.vase.poly.subset.complete$unpaired.disease <- jan2018.paired.vase.poly.subset.complete$sum.alive.d - #jan2018.paired.vase.poly.subset.complete$sum.pair.present #identifying pairs with vm-stragglers in disease

#jan2018.match1 <- merge(jan2018.ems, jan2018.paired.vase.poly.subset.complete, by = "paired.com.id")

#Identifying VIABLE Poly communities based on: must be 2 shoots alive in poly community that are represented in mono community.
jan2018.poly.2 <- jan2018 %>%
  dplyr::select(bin.id, paired.com.id, vase.id, loc.id, diversity.treatment, depth, family.id, color, alive) %>%
  filter(diversity.treatment == "poly") %>%
  filter(family.id == "S6" | family.id == "S3"| family.id == "S12"| family.id == "S11"| family.id == "S1" | family.id == "M9"| family.id == "M7"| family.id == "M6"| family.id == "M4"| family.id == "M3"| family.id == "M12"| family.id == "M10"| family.id == "D8"| family.id == "D6"| family.id == "D4"| family.id == "D3"| family.id == "D11"| family.id == "D10")

jan2018.poly.2.vase <- jan2018.poly.2 %>%
  group_by(bin.id, paired.com.id, vase.id, diversity.treatment, depth) %>%
  summarize(sum.alive = sum(alive)) %>%
  filter(sum.alive > 1)

View(jan2018.poly.2.vase)
  
#all alive shoots in target vases
jan2018.poly.2.alive <- jan2018 %>%
  dplyr::select(bin.id, paired.com.id, vase.id, loc.id, diversity.treatment, depth, family.id, color, alive) %>%
  filter(diversity.treatment == "poly") %>%
  filter(alive > 0) %>%
  filter(vase.id == 2 |vase.id == 6 |vase.id == 8 |vase.id == 9 |vase.id == 17|vase.id == 20|vase.id == 21|vase.id == 22|vase.id == 23|vase.id == 25|vase.id == 29|vase.id == 34|vase.id == 35|vase.id == 36|vase.id == 37|vase.id == 41|vase.id == 43|vase.id == 47|vase.id == 49|vase.id == 50|vase.id == 51|vase.id == 52|vase.id == 58|vase.id == 67|vase.id == 71|vase.id == 72|vase.id == 76|vase.id == 80|vase.id == 87|vase.id == 91|vase.id == 100|vase.id == 103|vase.id == 106|vase.id == 107|vase.id == 108|vase.id == 118)

#only alive shoots represented in mono from target vases
#jan2018.poly.2.alive.subset <- jan2018 %>%
   #dplyr::select(bin.id, paired.com.id, vase.id, loc.id, diversity.treatment, depth, family.id, color, alive) %>%
  #filter(diversity.treatment == "poly") %>%
  #filter(alive > 0) %>%
  #filter(family.id == "S6" | family.id == "S3"| family.id == "S12"| family.id == "S11"| family.id == "S1" | family.id == "M9"| family.id == "M7"| family.id == "M6"| family.id == "M4"| family.id == "M3"| family.id == "M12"| family.id == "M10"| family.id == "D8"| family.id == "D6"| family.id == "D4"| family.id == "D3"| family.id == "D11"| family.id == "D10") %>%
  #filter(vase.id == 2 |vase.id == 6 |vase.id == 8 |vase.id == 9 |vase.id == 17|vase.id == 20|vase.id == 21|vase.id == 22|vase.id == 23|vase.id == 25|vase.id == 29|vase.id == 34|vase.id == 35|vase.id == 36|vase.id == 37|vase.id == 41|vase.id == 43|vase.id == 47|vase.id == 49|vase.id == 50|vase.id == 51|vase.id == 52|vase.id == 58|vase.id == 67|vase.id == 71|vase.id == 72|vase.id == 76|vase.id == 80|vase.id == 87|vase.id == 91|vase.id == 100|vase.id == 103|vase.id == 106|vase.id == 107|vase.id == 108|vase.id == 118)

write.csv( jan2018.poly.2.alive, "PolyVases.csv" )

```

#Seagrass response
##Entire Experiment
```{r}
#jan2018.subset <- jan2018 %>%
 # filter(paired.com.id == "V93" | paired.com.id == "V91" | paired.com.id == "V107" | paired.com.id == "V95" | paired.com.id == "V55" | paired.com.id == "V71" | paired.com.id == "V59" | paired.com.id == "V25" | paired.com.id == "V21" | paired.com.id == "V23" |paired.com.id == "V17"| paired.com.id == "V47"| paired.com.id == "V49"| paired.com.id == "V11"| paired.com.id == "V37"| paired.com.id == "V53"| paired.com.id == "V89"| paired.com.id == "V77"| paired.com.id == "V41"| paired.com.id == "V43"| paired.com.id == "V15"| paired.com.id == "V83"| paired.com.id == "V39"| paired.com.id == "V5"| paired.com.id == "V79"| paired.com.id == "V87"| paired.com.id == "V81"| paired.com.id == "V1") %>%
#  group_by(bin.id, paired.com.id, diversity.treatment, disease.treatment, depth) %>%
#  summarize(sum.alive = sum(alive), mean.shoot.no = sum(shoot.no, na.rm = T)/sum(alive), mean.length.avg = sum(avg.length, na.rm = T)/sum(alive))

##Entire Experiment
#shoot number per vase
jan2018.fig <- jan2018 %>%
  group_by(vase.id, diversity.treatment, depth) %>%
  summarize(alive = sum(alive), shoot.no = sum(shoot.no), total.length = sum(total.length))

p <- ggplot(jan2018.fig, aes(x = diversity.treatment, y = shoot.no, fill = depth)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Diversity", y = "number of shoots per vase") +
  theme_classic()
p

bargraph.CI(x.factor = depth, response = shoot.no, group = diversity.treatment, data = jan2018.fig, ylab = "shoots per vase", legend = T, ylim = c(0,7))

model <-aov(shoot.no ~ depth*diversity.treatment, data = jan2018.fig)
anova(model)

#length per vase
p <- ggplot(jan2018.fig, aes(x = diversity.treatment, y = total.length, fill = depth)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Diversity", y = "total shoot length/vase") +
  theme_classic()
p

bargraph.CI(x.factor = depth, response = total.length, group = diversity.treatment, data = jan2018.fig, ylab = "total length per vase", legend = T, ylim = c(0,300))
model <-aov(total.length ~ depth*diversity.treatment, data = filter(jan2018.fig, diversity.treatment != "mix"))
anova(model)


#alive per vase
p <- ggplot(jan2018.fig, aes(x = diversity.treatment, y = alive, fill = depth)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Diversity", y = "proportion alive") +
  theme_classic()
p
bargraph.CI(x.factor = depth, response = alive, group = diversity.treatment, data = jan2018.fig, ylab = "alive per vase", legend = T, ylim = c(0,5))
model <-aov(alive ~ depth*diversity.treatment, data = filter(jan2018.fig, diversity.treatment != "mix"))
anova(model)

##shoot number per plant
p <- ggplot(jan2018.fig, aes(x = diversity.treatment, y = shoot.no/alive, fill = depth)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Diversity", y = "number of shoots per plant per vase") +
  theme_classic()
p
bargraph.CI(x.factor = depth, response = shoot.no/alive, group = diversity.treatment, data = jan2018.fig, ylab = "alive per vase", legend = T, ylim = c(0,5))
model <-aov((shoot.no/alive) ~ depth*diversity.treatment, data = filter(jan2018.fig, diversity.treatment != "mix"))
anova(model)
model <-aov((shoot.no/alive) ~ diversity.treatment, data = jan2018.fig)
anova(model)
```

##Focal Vases
###Data import and summary
```{r}
jan2018.focal <- read.csv("Vases_master.csv")

jan2018.focal <- jan2018.focal %>%
  group_by(vase.id, diversity.treatment, disease.treatment, depth) %>%
  summarize(alive = sum(alive), shoot.no = sum(shoot.no), total.length = sum(total.length.cm))
```

###shoots
```{r}
p <- ggplot(jan2018.focal, aes(x = diversity.treatment, y = shoot.no, fill = depth)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Diversity", y = "number of shoots per focal vase") +
  theme_classic()
p

bargraph.CI(x.factor = depth, response = shoot.no, group = diversity.treatment, data = jan2018.focal, ylab = "shoots per focal vase", legend = T, ylim = c(0,7))
model <-aov(shoot.no ~ depth*diversity.treatment, data = filter(jan2018.focal, diversity.treatment != "mix"))
anova(model)
model <-aov(shoot.no ~ diversity.treatment, data = jan2018.focal)
anova(model)
#disease * depth * diversity
bargraph.CI(x.factor = depth, response = shoot.no, group = disease.treatment, data = filter(jan2018.focal, diversity.treatment != "poly"), ylab = "shoots per focal vase", legend = T, ylim = c(0,10))
bargraph.CI(x.factor = depth, response = shoot.no, group = disease.treatment, data = filter(jan2018.focal, diversity.treatment != "mono"), ylab = "shoots per focal vase", legend = T, ylim = c(0,10))
#disease * diversity
bargraph.CI(x.factor = diversity.treatment, response = shoot.no, group = disease.treatment, data = jan2018.focal, ylab = "shoots per focal vase", legend = T, ylim = c(0,10))

#depth*disease model
model <-aov(shoot.no ~ depth*disease.treatment, data = jan2018.focal)
anova(model)
#diversity*disease model
model <-aov(shoot.no ~ diversity.treatment*disease.treatment, data = filter(jan2018.focal, depth != "mix"))
anova(model)
#depth*diversity*disease model
model.2 <-aov(shoot.no ~ diversity.treatment*disease.treatment*depth, data = filter(jan2018.focal, depth != "mix"))
anova(model.2)

jan2018.focal.filter <- filter(jan2018.focal, depth != "mix")

model <-aov(shoot.no ~ diversity.treatment*disease.treatment*depth, data = jan2018.focal)
anova(model)
```

###length
```{r}
p <- ggplot(jan2018.focal, aes(x = diversity.treatment, y = total.length, fill = depth)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Diversity", y = "total length per focal vase") +
  theme_classic()
p

bargraph.CI(x.factor = depth, response = total.length, group = diversity.treatment, data = jan2018.focal, ylab = "total length per focal vase", legend = T, ylim = c(0,300))
model <-aov(total.length ~ depth*diversity.treatment, data = filter(jan2018.focal, diversity.treatment != "mix"))
anova(model)
model <-aov(total.length ~ diversity.treatment, data = jan2018.focal)
anova(model)

#disease * depth
bargraph.CI(x.factor = depth, response = total.length, group = disease.treatment, data = jan2018.focal, ylab = "total length per focal vase", legend = T, ylim = c(0,400))
#disease * diversity
bargraph.CI(x.factor = diversity.treatment, response = total.length, group = disease.treatment, data = filter(jan2018.focal, diversity.treatment != "mix"), ylab = "total length per focal vase", legend = T, ylim = c(0,400))
#disease * depth * diversity
bargraph.CI(x.factor = depth, response = total.length, group = disease.treatment, data = filter(jan2018.focal, diversity.treatment != "poly"), ylab = "total length per focal vase", legend = T, ylim = c(0,400))
bargraph.CI(x.factor = depth, response = total.length, group = disease.treatment, data = filter(jan2018.focal, diversity.treatment != "mono"), ylab = "total length per focal vase", legend = T, ylim = c(0,400))

#depth*disease model
model <-aov(total.length ~ depth*disease.treatment, data = jan2018.focal)
anova(model)
#diversity*disease model
model <-aov(total.length ~ diversity.treatment*disease.treatment, data = filter(jan2018.focal, diversity.treatment != "mix"))
anova(model)
#depth*diversity*disease model
model.2 <-aov(total.length ~ diversity.treatment*disease.treatment*depth, data = filter(jan2018.focal, depth != "mix"))
anova(model.2)
```

###alive
```{r}
p <- ggplot(jan2018.focal, aes(x = diversity.treatment, y = alive, fill = depth)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Diversity", y = "alive per focal vase") +
  theme_classic()
p

bargraph.CI(x.factor = depth, response = alive, group = diversity.treatment, data = jan2018.focal, ylab = "plants alive per focal vase", legend = T, ylim = c(0,4))
model <-aov(alive ~ depth*diversity.treatment, data = filter(jan2018.focal, diversity.treatment != "mix"))
anova(model)
model <-aov(alive ~ diversity.treatment, data = jan2018.focal)
anova(model)

#disease * diversity
bargraph.CI(x.factor = diversity.treatment, response = alive, group = disease.treatment, data = filter(jan2018.focal, depth != "mix"), ylab = "alive", legend = T, ylim = c(0,4))
#disease * depth
bargraph.CI(x.factor = depth, response = alive, group = disease.treatment, data = jan2018.focal, ylab = "alive", legend = T, ylim = c(0,4))
#disease * depth * diversity
bargraph.CI(x.factor = depth, response = alive, group = disease.treatment, data = filter(jan2018.focal, diversity.treatment != "poly"), ylab = "alive", legend = T, ylim = c(0,4))
bargraph.CI(x.factor = depth, response = alive, group = disease.treatment, data = filter(jan2018.focal, diversity.treatment != "mono"), ylab = "alive", legend = T, ylim = c(0,4))

#depth*disease model
model <-aov(alive ~ depth*disease.treatment, data = jan2018.focal)
anova(model)
#diversity*disease model
model <-aov(alive ~ diversity.treatment*disease.treatment, data = filter(jan2018.focal, diversity.treatment != "mix"))
anova(model)
#depth*diversity*disease model
model.2 <-aov(alive ~ diversity.treatment*disease.treatment*depth, data = filter(jan2018.focal, depth != "mix"))
anova(model.2)
```

###shoots/plant per vase
```{r}
p <- ggplot(jan2018.focal, aes(x = diversity.treatment, y = shoot.no/alive, fill = depth)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Diversity", y = "number of shoots per plant per vase") +
  theme_classic()
p

bargraph.CI(x.factor = depth, response = shoot.no/alive, group = diversity.treatment, data = jan2018.focal, ylab = "shoots per plant per focal vase", legend = T, ylim = c(0,3))
model <-aov((shoot.no/alive) ~ depth*diversity.treatment, data = filter(jan2018.focal, diversity.treatment != "mix"))
anova(model)

#disease * diversity
bargraph.CI(x.factor = diversity.treatment, response = (shoot.no/alive), group = disease.treatment, data = filter(jan2018.focal, depth != "mix"), ylab = "shoots per plant per focal vase", legend = T, ylim = c(0,4))
#disease * depth
bargraph.CI(x.factor = depth, response = (shoot.no/alive), group = disease.treatment, data = jan2018.focal, ylab = "shoots per plant per focal vase", legend = T, ylim = c(0,4))
#disease * depth * diversity
bargraph.CI(x.factor = depth, response = (shoot.no/alive), group = disease.treatment, data = filter(jan2018.focal, diversity.treatment != "poly"), ylab = "shoots per plant per focal vase", legend = T, ylim = c(0,4))
bargraph.CI(x.factor = depth, response = (shoot.no/alive), group = disease.treatment, data = filter(jan2018.focal, diversity.treatment != "mono"), ylab = "shoots per plant per focal vase", legend = T, ylim = c(0,4))


#depth*disease model
model <-aov((shoot.no/alive) ~ depth*disease.treatment, data = jan2018.focal)
anova(model)
#diversity*disease model
model <-aov((shoot.no/alive) ~ diversity.treatment*disease.treatment, data = filter(jan2018.focal, diversity.treatment != "mix"))
anova(model)
#depth*diversity*disease model
model.2 <-aov((shoot.no/alive) ~ diversity.treatment*disease.treatment*depth, data = filter(jan2018.focal, depth != "mix"))
anova(model.2)
```

individual by shoots
```{r}
plot(shoot.no/alive ~ alive, data = jan2018.focal)
plot(shoot.no ~ alive, data = jan2018.focal)

jan2018.focal$alive.factor <- as.factor(jan2018.focal$alive)

p <- ggplot(jan2018.focal, aes(x = alive.factor, y = shoot.no/alive, fill = disease.treatment:diversity.treatment)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "alive", y = "number of shoots per plant") +
  theme_classic()
p

#the proportion of vectors per shoot is pretty consistent across shoots because the number of shoots per plant is consistent across treatments that vary in plant number (range: 1.5-2) resulting in a range in initial prevalence of 50-66%. The disease-poly treatment vases with 4 shoots have noticeably less shoots per plant than other treatments, but swapping without changing other factors is going to be very challenging without throwing off other comparisons which I think are more important. Also I think that if shoots/plant is averaged across all plant numbers then disease-poly treatment doesn't differ from other treatments in average shoots/plant
```

#inoculation analyses
##data import
```{r}
mar2018.inoc <- read.csv("Vases_master.csv")

str(mar2018.inoc)
```

##lesion success by shoot length
```{r}
plot(lesion.09.mar ~ terminal.shoot.length, data = filter(mar2018.inoc, disease.treatment == "disease"))

model <- glm(cbind(lesion.09.mar,(1-lesion.09.mar)) ~ terminal.shoot.length, family = binomial(link = 'logit'), data = filter(mar2018.inoc, disease.treatment == "disease"))#logistic regression
summary(model)

curve(predict(model, data.frame(terminal.shoot.length = x), type = "resp"), add = T)
#not significant and not much of a trend
```

##family identity
```{r}
p <- ggplot(filter(mar2018.1month, disease.treatment == "disease" & depth == "deep"), aes(x = family.id, y = lesion.present.26MAR, fill = diversity.treatment)) + 
  scale_fill_manual(values = c("white", "gray")) +
  #geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 1) +
  #stat_summary(fun.y = mean, geom = "point", size = 2, color = "red", position = position_dodge(1)) +
  labs(x = "family ID", y = "proportion of successful infections") +
  theme_classic()
p
```

##Randall model
```{r}
#Maybe we run a model that tests family*diversity on inoculation success (as a binomial)? I don’t think diversity will be significant, but probably good to have it in the model. We could also see whether depth*diversity, or depth(family)*diversity show anything.

mar2018.inoc.exclude <- subset(mar2018.inoc, family.id != "D9")
mar2018.inoc.exclude <- subset(mar2018.inoc.exclude, family.id != "M8")
mar2018.inoc.exclude <- subset(mar2018.inoc.exclude, family.id != "S10")
mar2018.inoc.exclude <- subset(mar2018.inoc.exclude, family.id != "S4")
mar2018.inoc.exclude <- subset(mar2018.inoc.exclude, family.id != "S5")
mar2018.inoc.exclude <- subset(mar2018.inoc.exclude, family.id != "D1")
mar2018.inoc.exclude <- subset(mar2018.inoc.exclude, family.id != "D2")
mar2018.inoc.exclude <- subset(mar2018.inoc.exclude, family.id != "M11")

model <- glm(cbind(lesion.09.mar,(1-lesion.09.mar)) ~ depth*diversity.treatment, family = binomial(link = 'logit'), data = filter(mar2018.inoc, disease.treatment == "disease"))#logistic regression
Anova(model)
anova(model, test = "LR")

```

#1MONTH Analyses

##data import
```{r}
mar2018.1month <- read.csv("Vases_master_1month.csv")

str(mar2018.1month)

mar2018.1month$shoot.no.pc <-
  (mar2018.1month$shoot.no.26MAR-mar2018.1month$shoot.no.24JAN)/mar2018.1month$shoot.no.24JAN #shoot.no change per shoot per plant
mar2018.1month$total.length.pc <- 
  (mar2018.1month$total.length.26MAR/mar2018.1month$shoot.no.26MAR-mar2018.1month$total.length.24JAN/mar2018.1month$shoot.no.24JAN)#/(mar2018.1month$total.length.24JAN/mar2018.1month$shoot.no.24JAN) #avg length change per initial avg. shoot length per plant
mar2018.1month$infected.shoot.length.pc <- 
  (mar2018.1month$infected.shoot.length.26MAR-mar2018.1month$infected.shoot.length.24JAN)/mar2018.1month$infected.shoot.length.24JAN #infected shoot length change per initial infected shoot length per plant


mar2018.1month.vase <- mar2018.1month %>%
  group_by(diversity.treatment, disease.treatment, vase.id) %>%
  summarize(shoot.no.pc.vase = mean(shoot.no.pc, na.rm = T),
            total.length.pc.vase = mean(total.length.pc, na.rm = T),
            infected.shoot.length.pc.vase = mean(infected.shoot.length.pc, na.rm = T),
            lesion.spread.26MAR.vase = mean(lesion.spread.26MAR, na.rm = T))

mar2018.1month.vase.NM <- mar2018.1month %>% #NO MIX
  filter(depth != "mix") %>%
  group_by(diversity.treatment, disease.treatment, depth, vase.id) %>%
  summarize(shoot.no.pc.vase = mean(shoot.no.pc, na.rm = T),
            total.length.pc.vase = mean(total.length.pc, na.rm = T),
            infected.shoot.length.pc.vase = mean(infected.shoot.length.pc, na.rm = T),
            lesion.spread.26MAR.vase = mean(lesion.spread.26MAR, na.rm = T))
```

##ALL Figures
```{r}
p <- ggplot(mar2018.1month.vase, aes(x = disease.treatment, y = shoot.no.pc.vase, fill = diversity.treatment)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Disease", y = "Change in number of shoots per shoot per plant per vase") +
  theme_classic()
p

bargraph.CI(x.factor = disease.treatment, response = shoot.no.pc.vase, group = diversity.treatment, data = mar2018.1month.vase., ylab = "Proportion change in number of shoots per plant per vase", legend = T)

p <- ggplot(mar2018.1month.vase, aes(x = disease.treatment, y = total.length.pc.vase, fill = diversity.treatment)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Disease", y = "Proportional change in total length per plant per vase") +
  theme_classic()
p

bargraph.CI(x.factor = disease.treatment, response = total.length.pc.vase, group = diversity.treatment, data = mar2018.1month.vase, ylab = "Proportional change in total length per plant per vase", legend = T)

p <- ggplot(mar2018.1month.vase, aes(x = disease.treatment, y = infected.shoot.length.pc.vase , fill = diversity.treatment)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Disease", y = "Proportion change in length of inoculated shoots per plant per vase") +
  theme_classic()
p

bargraph.CI(x.factor = disease.treatment, response = infected.shoot.length.pc.vase, group = diversity.treatment, data = mar2018.1month.vase, ylab = "Proportion change in length of inoculated shoots per plant per vase", legend = T)


p <- ggplot(filter(mar2018.1month.vase, disease.treatment == "disease"), aes(x = diversity.treatment, y = lesion.spread.26MAR.vase, fill = diversity.treatment)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 0.5) +
  stat_summary(fun.y = mean, geom = "point", size = 3, color = "red", position = position_dodge(1)) +
  labs(x = "Diversity", y = "Delta total length per shoot per vase") +
  theme_classic()
p

bargraph.CI(x.factor = diversity.treatment, response = lesion.spread.26MAR.vase, data = filter(mar2018.1month.vase, disease.treatment == "disease"), ylab = "Mean Lesion Length per vase", legend = T)

p <- ggplot(filter(mar2018.1month, disease.treatment == "disease" & depth == "deep"), aes(x = family.id, y = lesion.spread.26MAR)) + 
  #scale_fill_manual(values = c("white", "gray")) +
  #geom_violin(position = position_dodge(1)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', position = position_dodge(1), dotsize  = 1) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "red", position = position_dodge(1)) +
  labs(x = "family ID", y = "Lesion Length (cm)") +
  ylim(0,30) +
  theme_classic()
p
```

##Analyses
###Shoot Number
```{r}
#Shoot Number MODEL 

#model.shoot.no.1month <-aov(shoot.no.pc.vase ~ diversity.treatment*disease.treatment, data = mar2018.1month.vase)
#anova(model.shoot.no.1month)

#####no poly mix treatment
model.shoot.no.1month <- aov(shoot.no.26MAR ~ diversity.treatment*disease.treatment*depth + shoot.no.24JAN, data = filter(mar2018.1month, depth != "mix"))
anova(model.shoot.no.1month)

shoot.no.regression <- lm(shoot.no.26MAR ~ shoot.no.24JAN, data = filter(mar2018.1month, depth != "mix"))
summary(shoot.no.regression)
plot(shoot.no.26MAR ~ shoot.no.24JAN, data = filter(mar2018.1month, depth != "mix"), xlim = c(0,6), ylim = c(0,6))
abline(shoot.no.regression)

mar2018.1month$shoot.no.resid <- mar2018.1month$shoot.no.26MAR - (mar2018.1month$shoot.no.24JAN*1.13332 + -0.01156)

#posthoc <- TukeyHSD(x=model.shoot.no.1month, 'diversity.treatment:depth', conf.level=0.95)
#posthoc
#multcompLetters(posthoc$'diversity.treatment:depth'[, "p adj"])

#bargraph.CI(x.factor = family.id, response = shoot.no.resid, data = filter(mar2018.1month, depth != "mix"), ylim = c(-1.5,1))

bargraph.CI(x.factor = depth, response = shoot.no.resid, group = diversity.treatment, data = filter(mar2018.1month, depth != "mix"), ylim = c(-1,1), legend = T)

bargraph.CI(x.factor = disease.treatment, response = shoot.no.resid, data = filter(mar2018.1month, depth != "mix"), ylim = c(-0.25,0.25))
```

###Average Length
```{r}
#model.total.length.1month <- aov(total.length.pc.vase ~ diversity.treatment*disease.treatment, data = mar2018.1month.vase )
#anova(model.total.length.1month)

#model.total.length.1month.mono <- aov(total.length.1month.vase ~ disease.treatment, data = filter(mar2018.1month.vase, diversity.treatment == "mono"))
#anova(model.total.length.1month.mono)
#model.total.length.1month.poly <- aov(total.length.1month.vase ~ disease.treatment, data = filter(mar2018.1month.vase, diversity.treatment == "poly"))
#anova(model.total.length.1month.poly)

model.total.length.1month <- aov(total.length.26MAR ~ diversity.treatment*disease.treatment*depth + shoot.no.24JAN + total.length.24JAN, data = filter(mar2018.1month, depth != "mix"))
anova(model.total.length.1month)

bargraph.CI(x.factor = disease.treatment, response = total.length.pc, data = filter(mar2018.1month, depth != "mix"), ylim = c(-50,0))
```

###Infected shoot length
```{r}
model.infected.shoot.length.1month <- aov(infected.shoot.length.26MAR ~ diversity.treatment*disease.treatment*depth + infected.shoot.length.24JAN, data = filter( mar2018.1month, depth != "mix"))
anova(model.infected.shoot.length.1month)

infected.shoot.length.regression <- lm(infected.shoot.length.26MAR ~ infected.shoot.length.24JAN, data = filter(mar2018.1month, depth != "mix"))
summary(infected.shoot.length.regression)
plot(infected.shoot.length.26MAR ~ infected.shoot.length.24JAN, data = filter(mar2018.1month, depth != "mix"), xlim = c(0,150), ylim = c(0,150))
abline(infected.shoot.length.regression)

mar2018.1month$infected.shoot.length.resid <- mar2018.1month$infected.shoot.length.26MAR - (mar2018.1month$infected.shoot.length.24JAN*0.87745 + 0.09082)

bargraph.CI(x.factor = disease.treatment, response = infected.shoot.length.resid, data = filter(mar2018.1month, depth != "mix"), ylim = c(-3,3))

bargraph.CI(x.factor = depth, response = infected.shoot.length.resid, data = filter(mar2018.1month, depth != "mix"), ylim = c(-4,3))
```

###Lesion length
```{r}
model.lesion.length <- aov(lesion.spread.26MAR.vase ~ diversity.treatment, data = filter(mar2018.1month.vase, disease.treatment == "disease"))
anova(model.lesion.length)
```

##No Mix Figures
```{r}
#switch up filter depth
bargraph.CI(x.factor = disease.treatment, response = shoot.no.pc.vase, group = diversity.treatment, data = filter(mar2018.1month.vase.NM, depth == "deep"), ylab = "Proportion change in number of shoots per plant per vase", legend = T)

#
bargraph.CI(x.factor = disease.treatment, response = total.length.pc.vase, group = diversity.treatment, data = filter(mar2018.1month.vase.NM, depth == "deep"), ylab = "Proportional change in total length per plant per vase", legend = T)

#
bargraph.CI(x.factor = disease.treatment, response = infected.shoot.length.pc.vase, group = diversity.treatment, data = filter(mar2018.1month.vase.NM, depth == "deep"), ylab = "Proportion change in length of inoculated shoots per plant per vase", legend = T)

#
bargraph.CI(x.factor = diversity.treatment, response = lesion.spread.26MAR.vase, group = depth, data = filter(mar2018.1month.vase.NM, disease.treatment == "disease"), ylab = "Mean Lesion Length per vase", legend = T)
```

#1MONTH Analyses for Team Seagrass

##GENOTYPE ID ANALYSES
###manipulating data
```{r}
#all data manipulation done in excel
  #randomly selected 1 shoot from all vases when more than 1 shoot per family per vase (i.e. monocultures)
  #exculded all shoots not present at initial sampling (24JAN)
d.GID <- read.csv("d.GID.csv")

#Exculding families with only 1 shoot
d.GID <- filter(d.GID, family.id != "M8")
d.GID <- filter(d.GID, family.id != "S4")

#delta responses
d.GID$delta.shoots <- d.GID$shoot.no.26MAR-d.GID$shoot.no.24JAN
d.GID$delta.infected.shoot <- d.GID$infected.shoot.length.26MAR-d.GID$infected.shoot.length.24JAN

```

###analyses and figures
```{r}
#Shoot Number
model.GID <- aov(shoot.no.26MAR ~ family.id + shoot.no.24JAN, data = d.GID)
anova(model.GID)

p <- ggplot(d.GID, aes(x = family.id, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = family.id)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = family.id), size = 0.5) +
  labs(x = "Family ID", y = "Delta Shoot No.") +
  theme_classic()
p

#Infected Shoot Length
model.GID <- aov(infected.shoot.length.26MAR ~ family.id + infected.shoot.length.24JAN, data = d.GID)
anova(model.GID)

p <- ggplot(d.GID, aes(x = family.id, y = delta.infected.shoot)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = family.id)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = family.id), size = 0.5) +
  labs(x = "Family ID", y = "Delta Shoot No.") +
  theme_classic()
p
```

##MONO_POLY ANALYSES
###manipulating data
```{r}
  #Reading in data (No changes from 1 month vase master data)
d.MP <- read.csv("d.MonoPoly.csv")

  #Excluding all dead plants at initial sampling, all plants in mix depth polycultures, all plants from families that are not represented in at least 1 of each of 4 diversity*disease combinations
  #Then taking mean number of shoots for each family in each of 4 diversity by disease combinations
d.MP <- d.MP %>%
  filter(alive.24JAN > 0) %>%
  filter(depth != "mix") %>%
  filter(family.id != "D1") %>%
  filter(family.id != "D2") %>%
  filter(family.id != "D9") %>%
  filter(family.id != "M10") %>%
  filter(family.id != "M11") %>%
  filter(family.id != "M8") %>%
  filter(family.id != "S10") %>%
  filter(family.id != "S4") %>%
  filter(family.id != "S5") %>%
  group_by(diversity.treatment, disease.treatment, depth, family.id) %>%
  summarize(shoots.initial = mean(shoot.no.24JAN, na.rm = T), shoots.final = mean(shoot.no.26MAR, na.rm = T), plants.initial = sum(alive.24JAN, na.rm = T), isl.initial = mean(infected.shoot.length.24JAN, na.rm = T), isl.final = mean(infected.shoot.length.26MAR))

#delta responses
d.MP$delta.shoots <- d.MP$shoots.final - d.MP$shoots.initial
d.MP$delta.isl <- d.MP$isl.final - d.MP$isl.initial
```

###analyses and figures
```{r}
#Shoot Number Family Identity
model.MP <- aov(shoots.final ~ depth/family.id + shoots.initial, data = d.MP)
anova(model.MP)

#Shoot Number Mono-Poly
model.MP <- aov(shoots.final ~ diversity.treatment*depth*disease.treatment + shoots.initial, data = d.MP)
anova(model.MP)

p <- ggplot(d.MP, aes(x = depth, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = depth), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = depth), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Depth", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,1,0.2), limits = c(0,1)) +
  coord_cartesian(ylim = c(0,0.5)) +
  theme_classic()
p

p <- ggplot(d.MP, aes(x = family.id, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = depth), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = depth), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Family ID", y = "Delta Shoot No.") +
  theme_classic()
p

#Infected Shoot Length
model.MP <- aov(isl.final ~ diversity.treatment*depth*disease.treatment + isl.initial + plants.initial, data = d.MP)
anova(model.MP)

p <- ggplot(d.MP, aes(x = disease.treatment, y = delta.isl)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = depth)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = depth), size = 0.5) +
  labs(x = "Depth", y = "Delta Infected Shoot Length") +
  theme_classic()
p
```

##EXP_OBS ANALYSES
###manipulating data
```{r}
#Calculating expected values
d.OE.expected <- read.csv("d.MonoPoly.csv")
d.OE.expected <- d.OE.expected %>%
  filter(alive.24JAN > 0) %>%
  filter(depth != "mix") %>%
  group_by(diversity.treatment, disease.treatment, depth, family.id) %>%
  summarize(shoots.initial = mean(shoot.no.24JAN, na.rm = T), shoots.final = mean(shoot.no.26MAR, na.rm = T), plants.initial = sum(alive.24JAN, na.rm = T), isl.initial = mean(infected.shoot.length.24JAN, na.rm = T), isl.final = mean(infected.shoot.length.26MAR))

d.OE.expected.total <- d.OE.expected %>%
  group_by(diversity.treatment, disease.treatment) %>%
  summarize(shoots.initial = mean(shoots.initial), shoots.final = mean(shoots.final))

#Uploading data with expected values for shoot number for each shoot (entered in excel using data produced above as reference)
d.OE <- read.csv("d.OE.csv")

#summing shoot values within each vase and removing vases where shoots do not have monoculture to calculate expected value
d.OE <- d.OE %>%
  group_by(diversity.treatment, disease.treatment, depth, vase.id) %>%
  summarize(obs.sn.initial = sum(obs.sn.initial), obs.sn.final = sum(obs.sn.final), exp.sn.initial = sum(exp.sn.initial), exp.sn.final = sum(exp.sn.final)) %>%
  filter(exp.sn.initial != "NA")

#calculating change in shoots expected and observed at vase level
d.OE$delta.shoots.exp <- d.OE$exp.sn.final-d.OE$exp.sn.initial
d.OE$delta.shoots.obs <- d.OE$obs.sn.final-d.OE$obs.sn.initial

#creating single columns for observed and expected values
#aka creating column that seperates observed and expected values
d.OE.A <- subset(d.OE, select = c(diversity.treatment:obs.sn.initial, exp.sn.initial))
d.OE.A <- gather(d.OE.A, obs.exp.initial, sn.initial, obs.sn.initial:exp.sn.initial)
d.OE.A$exp.obs <- d.OE.A$obs.exp.initial
d.OE.A$exp.obs[d.OE.A$exp.obs=="obs.sn.initial"] <- "obs"
d.OE.A$exp.obs[d.OE.A$exp.obs=="exp.sn.initial"] <- "exp"
d.OE.A <- subset(d.OE.A, select = c(diversity.treatment:vase.id, sn.initial:exp.obs))

d.OE.B <- subset(d.OE, select = c(diversity.treatment:vase.id, obs.sn.final, exp.sn.final))
d.OE.B<- gather(d.OE.B, obs.exp.final, sn.final, obs.sn.final:exp.sn.final)
d.OE.B$exp.obs <- d.OE.B$obs.exp.final
d.OE.B$exp.obs[d.OE.B$exp.obs=="obs.sn.final"] <- "obs"
d.OE.B$exp.obs[d.OE.B$exp.obs=="exp.sn.final"] <- "exp"
d.OE.B <- subset(d.OE.B, select = c(diversity.treatment:vase.id, sn.final:exp.obs))

d.OE.C <- subset(d.OE, select = c(diversity.treatment:vase.id, delta.shoots.obs, delta.shoots.exp))
d.OE.C <- gather(d.OE.C, exp.obs, delta.shoots, delta.shoots.obs:delta.shoots.exp)
d.OE.C$exp.obs[d.OE.C$exp.obs=="delta.shoots.obs"] <- "obs"
d.OE.C$exp.obs[d.OE.C$exp.obs=="delta.shoots.exp"] <- "exp"
d.OE.C <- subset(d.OE.C, select = c(diversity.treatment:vase.id, delta.shoots:exp.obs))

d.OE <- merge(d.OE.A, d.OE.B, by.all = c("diversity.treatment", "disease.treatment", "depth", "vase.id", "exp.obs"))

d.OE <- merge(d.OE, d.OE.C, by.all = c("diversity.treatment", "disease.treatment", "depth", "vase.id", "exp.obs"))
```

###analyses and figures
```{r}
#Shoot Number
model.OE <- aov(sn.final ~ exp.obs + exp.obs:disease.treatment + exp.obs:depth + exp.obs:disease.treatment:depth + sn.initial, data = d.OE)
anova(model.OE) #Do I need to pair vases??

p <- ggplot(d.OE, aes(x = exp.obs, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", aes(colour = depth), size = 2) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = depth), size = 0.5) +
  labs(x = "Expected vs Observed", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,3,0.5)) +
  coord_cartesian(ylim = c(-1,3)) +
  theme_classic()
p

p <- ggplot(filter(d.OE, disease.treatment == "disease"), aes(x = exp.obs, y = delta.shoots)) + 
  stat_summary(fun.y = "mean", geom = "point", aes(colour = depth), size = 2) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", aes(colour = depth), size = 0.5) +
  labs(x = "Expected vs Observed", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(-1,3,0.5)) +
  coord_cartesian(ylim = c(-1,3)) +
  theme_classic()
p

```

#2MONTH Analyses for Team Seagrass

##MONO_POLY ANALYSES
###manipulating data
```{r}
  #Reading in data from 2 month vase master data
d.MP.2month <- read.csv("Vases_master_2month.csv")

  #Excluding all dead plants at initial sampling, all plants in mix depth polycultures, all plants from families that are not represented in at least 1 of each of 4 diversity*disease combinations
  #Then taking mean number of shoots for each family in each of 4 diversity by disease combinations
d.MP.2month <- d.MP.2month %>%
  filter(alive.24JAN > 0) %>%
  filter(depth != "mix") %>%
  filter(family.id != "D1") %>%
  filter(family.id != "D2") %>%
  filter(family.id != "D9") %>%
  filter(family.id != "M10") %>%
  filter(family.id != "M11") %>%
  filter(family.id != "M8") %>%
  filter(family.id != "S10") %>%
  filter(family.id != "S4") %>%
  filter(family.id != "S5") %>%
  group_by(diversity.treatment, disease.treatment, depth, family.id) %>%
  summarize(shoots.initial = mean(shoot.no.24JAN, na.rm = T), shoots.final = mean(shoot.no.24APR, na.rm = T), plants.initial = sum(alive.24APR, na.rm = T), isl.initial = mean(infected.shoot.length.24JAN, na.rm = T), isl.final = mean(infected.shoot.length.24APR))

#delta responses
d.MP.2month$delta.shoots <- d.MP.2month$shoots.final - d.MP.2month$shoots.initial
d.MP.2month$delta.isl <- d.MP.2month$isl.final - d.MP.2month$isl.initial
```

###analyses and figures
```{r}
#Shoot Number Family Identity
model.MP <- aov(shoots.final ~ depth/family.id + shoots.initial, data = d.MP.2month)
anova(model.MP)

#Shoot Number Mono-Poly
model.MP <- aov(shoots.final ~ diversity.treatment*depth*disease.treatment + shoots.initial, data = d.MP.2month)
anova(model.MP)

#depth:disease.treatment (P = 0.038)
p <- ggplot(d.MP.2month, aes(x = disease.treatment, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = depth), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = depth), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Depth", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,1,0.2), limits = c(0,1)) +
  coord_cartesian(ylim = c(-0.5,1)) +
  theme_classic()
p

#diversity.treatment:disease.treatment (P = 0.078)
p <- ggplot(d.MP.2month, aes(x = disease.treatment, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = diversity.treatment), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = diversity.treatment), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Depth", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,1,0.2), limits = c(0,1)) +
  coord_cartesian(ylim = c(-0.5,1)) +
  theme_classic()
p

#depth:family.id (P = 0.058)
p <- ggplot(d.MP.2month, aes(x = family.id, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = depth), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = depth), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Family ID", y = "Delta Shoot No.") +
  theme_classic()
p
```

##EXP_OBS ANALYSES
###manipulating data
```{r}
#Calculating expected values
d.OE.expected.2month <- read.csv("Vases_master_2month.csv")
d.OE.expected.2month <- d.OE.expected.2month %>%
  filter(alive.24JAN > 0) %>%
  filter(depth != "mix") %>%
  group_by(diversity.treatment, disease.treatment, depth, family.id) %>%
  summarize(shoots.initial = mean(shoot.no.24JAN, na.rm = T), shoots.final = mean(shoot.no.24APR, na.rm = T), plants.initial = sum(alive.24JAN, na.rm = T), isl.initial = mean(infected.shoot.length.24JAN, na.rm = T), isl.final = mean(infected.shoot.length.24APR))

d.OE.expected.2month.total <- d.OE.expected.2month %>%
  group_by(diversity.treatment, disease.treatment) %>%
  summarize(shoots.initial = mean(shoots.initial), shoots.final = mean(shoots.final))

#Uploading data with expected values for shoot number for each shoot (entered in excel using data produced above as reference)
d.OE.2month <- read.csv("d.OE.2month.csv")

#summing shoot values within each vase and removing vases where shoots do not have monoculture to calculate expected value
d.OE.2month <- d.OE.2month %>%
  group_by(diversity.treatment, disease.treatment, depth, vase.id) %>%
  summarize(obs.sn.initial = sum(obs.sn.initial), obs.sn.final = sum(obs.sn.2month), exp.sn.initial = sum(exp.sn.initial), exp.sn.final = sum(exp.sn.2month)) %>%
  filter(exp.sn.initial != "NA")

#calculating change in shoots expected and observed at vase level
d.OE.2month$delta.shoots.exp <- d.OE.2month$exp.sn.final - d.OE.2month$exp.sn.initial
d.OE.2month$delta.shoots.obs <- d.OE.2month$obs.sn.final - d.OE.2month$obs.sn.initial

#creating single columns for observed and expected values
#aka creating column that seperates observed and expected values
d.OE.2month.A <- subset(d.OE.2month, select = c(diversity.treatment:obs.sn.initial, exp.sn.initial))
d.OE.2month.A <- gather(d.OE.2month.A, obs.exp.initial, sn.initial, obs.sn.initial:exp.sn.initial)
d.OE.2month.A$exp.obs <- d.OE.2month.A$obs.exp.initial
d.OE.2month.A$exp.obs[d.OE.2month.A$exp.obs=="obs.sn.initial"] <- "obs"
d.OE.2month.A$exp.obs[d.OE.2month.A$exp.obs=="exp.sn.initial"] <- "exp"
d.OE.2month.A <- subset(d.OE.2month.A, select = c(diversity.treatment:vase.id, sn.initial:exp.obs))

d.OE.2month.B <- subset(d.OE.2month, select = c(diversity.treatment:vase.id, obs.sn.final, exp.sn.final))
d.OE.2month.B<- gather(d.OE.2month.B, obs.exp.final, sn.final, obs.sn.final:exp.sn.final)
d.OE.2month.B$exp.obs <- d.OE.2month.B$obs.exp.final
d.OE.2month.B$exp.obs[d.OE.2month.B$exp.obs=="obs.sn.final"] <- "obs"
d.OE.2month.B$exp.obs[d.OE.2month.B$exp.obs=="exp.sn.final"] <- "exp"
d.OE.2month.B <- subset(d.OE.2month.B, select = c(diversity.treatment:vase.id, sn.final:exp.obs))

d.OE.2month.C <- subset(d.OE.2month, select = c(diversity.treatment:vase.id, delta.shoots.obs, delta.shoots.exp))
d.OE.2month.C <- gather(d.OE.2month.C, exp.obs, delta.shoots, delta.shoots.obs:delta.shoots.exp)
d.OE.2month.C$exp.obs[d.OE.2month.C$exp.obs=="delta.shoots.obs"] <- "obs"
d.OE.2month.C$exp.obs[d.OE.2month.C$exp.obs=="delta.shoots.exp"] <- "exp"
d.OE.2month.C <- subset(d.OE.2month.C, select = c(diversity.treatment:vase.id, delta.shoots:exp.obs))

d.OE.2month <- merge(d.OE.2month.A, d.OE.2month.B, by.all = c("diversity.treatment", "disease.treatment", "depth", "vase.id", "exp.obs"))

d.OE.2month <- merge(d.OE.2month, d.OE.2month.C, by.all = c("diversity.treatment", "disease.treatment", "depth", "vase.id", "exp.obs"))
```

###analyses and figures
```{r}
#Shoot Number
model.OE <- aov(sn.final ~ exp.obs + exp.obs:disease.treatment + exp.obs:depth + exp.obs:disease.treatment:depth + sn.initial, data = d.OE.2month)
anova(model.OE) #Do I need to pair vases??

p <- ggplot(d.OE, aes(x = exp.obs, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", aes(colour = depth), size = 2) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = depth), size = 0.5) +
  labs(x = "Expected vs Observed", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,3,0.5)) +
  coord_cartesian(ylim = c(-1,3)) +
  theme_classic()
p

p <- ggplot(filter(d.OE, disease.treatment == "control"), aes(x = exp.obs, y = delta.shoots)) + 
  stat_summary(fun.y = "mean", geom = "point", aes(colour = depth), size = 2) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", aes(colour = depth), size = 0.5) +
  labs(x = "Expected vs Observed", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(-1,3,0.5)) +
  coord_cartesian(ylim = c(-1,3)) +
  theme_classic()
p

```

##VASE LEVEL ANALYSES
###manipulating data
```{r}
d.vase.2month <- read.csv("Vases_master_2month.csv")

d.vase.2month <- d.vase.2month %>%
  filter(alive.24JAN > 0) %>%
  filter(depth != "mix") %>%
  group_by(diversity.treatment, disease.treatment, depth, paired.com.id.final) %>%
  summarize(shoots.initial = sum(shoot.no.24JAN, na.rm = T), shoots.final = sum(shoot.no.24APR, na.rm = T), plants.initial = sum(alive.24JAN, na.rm = T), tl.initial = sum(total.length.24JAN, na.rm = T), tl.final = sum(total.length.24APR, na.rm = T), isl.initial = mean(infected.shoot.length.24JAN, na.rm = T), isl.final = mean(infected.shoot.length.24APR, na.rm = T))

#delta responses
d.vase.2month$delta.shoots <- d.vase.2month$shoots.final - d.vase.2month$shoots.initial
d.vase.2month$delta.tl <- d.vase.2month$tl.final - d.vase.2month$tl.initial
```

###analyses and figures
```{r}
#shoot number
model.vase <- aov(shoots.final ~ diversity.treatment*depth*disease.treatment + shoots.initial, data = d.vase.2month)
anova(model.vase)

#diversity.treatment (P = 0.093)
p <- ggplot(d.vase.2month, aes(x = diversity.treatment, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = diversity.treatment), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = diversity.treatment), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Diversity Treatment", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,1,0.2), limits = c(0,1)) +
  coord_cartesian(ylim = c(0,2)) +
  theme_classic()
p

#disease.treatment (P = 0.0097)
p <- ggplot(d.vase.2month, aes(x = disease.treatment, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = disease.treatment), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = disease.treatment), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Disease Treatment", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,1,0.2), limits = c(0,1)) +
  coord_cartesian(ylim = c(0,2)) +
  theme_classic()
p

#diversity.treatment:depth (P = 0.045)
p <- ggplot(d.vase.2month, aes(x = diversity.treatment, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = depth), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = depth), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Diversity Treatment", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,1,0.2), limits = c(0,1)) +
  coord_cartesian(ylim = c(-0.5,3)) +
  theme_classic()
p

#disease.treatment:depth (P = 0.086)
p <- ggplot(d.vase.2month, aes(x = disease.treatment, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = depth), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = depth), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Disease Treatment", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,1,0.2), limits = c(0,1)) +
  coord_cartesian(ylim = c(-1,3)) +
  theme_classic()
p

#diversity.treatment:disease.treatment:depth (P = 0.072)
p <- ggplot(filter(d.vase.2month, depth == "shallow"), aes(x = disease.treatment, y = delta.shoots)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = diversity.treatment), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = diversity.treatment), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Disease Treatment", y = "Delta Shoot No.") +
  #scale_y_continuous(breaks=seq(0,1,0.2), limits = c(0,1)) +
  coord_cartesian(ylim = c(-1,5)) +
  theme_classic()
p

#Total length
model.vase <- aov(tl.final ~ diversity.treatment*disease.treatment + tl.initial, data = filter(d.vase.2month, depth == "shallow"))
anova(model.vase)

#diversity.treatment:disease.treatment:depth (P = 0.037)
p <- ggplot(filter(d.vase.2month, depth == "shallow"), aes(x = disease.treatment, y = delta.tl)) + 
  stat_summary(fun.y = mean, geom = "point", size = 2, aes(colour = diversity.treatment), position = position_dodge(0.5)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(colour = diversity.treatment), position = position_dodge(0.5), size = 0.5) +
  labs(x = "Disease Treatment", y = "Delta Total Length (cm)") +
  #scale_y_continuous(breaks=seq(0,1,0.2), limits = c(0,1)) +
  coord_cartesian(ylim = c(-200,100)) +
  theme_classic()
p

```

